---
title: "Class19"
author: "Daniel Xu"
format: pdf
---

Q1)

```{r}

cdc<-data.frame(
                          Year = c(1922L,
                                   1923L,1924L,1925L,1926L,1927L,1928L,
                                   1929L,1930L,1931L,1932L,1933L,1934L,1935L,
                                   1936L,1937L,1938L,1939L,1940L,1941L,
                                   1942L,1943L,1944L,1945L,1946L,1947L,1948L,
                                   1949L,1950L,1951L,1952L,1953L,1954L,
                                   1955L,1956L,1957L,1958L,1959L,1960L,
                                   1961L,1962L,1963L,1964L,1965L,1966L,1967L,
                                   1968L,1969L,1970L,1971L,1972L,1973L,
                                   1974L,1975L,1976L,1977L,1978L,1979L,1980L,
                                   1981L,1982L,1983L,1984L,1985L,1986L,
                                   1987L,1988L,1989L,1990L,1991L,1992L,1993L,
                                   1994L,1995L,1996L,1997L,1998L,1999L,
                                   2000L,2001L,2002L,2003L,2004L,2005L,
                                   2006L,2007L,2008L,2009L,2010L,2011L,2012L,
                                   2013L,2014L,2015L,2016L,2017L,2018L,
                                   2019L),
  No..Reported.Pertussis.Cases = c(107473,
                                   164191,165418,152003,202210,181411,
                                   161799,197371,166914,172559,215343,179135,
                                   265269,180518,147237,214652,227319,103188,
                                   183866,222202,191383,191890,109873,
                                   133792,109860,156517,74715,69479,120718,
                                   68687,45030,37129,60886,62786,31732,28295,
                                   32148,40005,14809,11468,17749,17135,
                                   13005,6799,7717,9718,4810,3285,4249,
                                   3036,3287,1759,2402,1738,1010,2177,2063,
                                   1623,1730,1248,1895,2463,2276,3589,
                                   4195,2823,3450,4157,4570,2719,4083,6586,
                                   4617,5137,7796,6564,7405,7298,7867,
                                   7580,9771,11647,25827,25616,15632,10454,
                                   13278,16858,27550,18719,48277,28639,
                                   32971,20762,17972,18975,15609,18617)
)
```

```{r}
library(ggplot2)
ggplot(cdc,aes(Year,No..Reported.Pertussis.Cases))+geom_point()+geom_line()+labs(title="Number of Reported Pertussis Cases over Time")
oplot<-ggplot(cdc,aes(Year,No..Reported.Pertussis.Cases))+geom_point()+geom_line()+labs(title="Number of Reported Pertussis Cases over Time")
```

```{r}
oplot+geom_vline(xintercept=1946,color="red",linetype=4)+geom_vline(xintercept=1996,color="blue",linetype=4)+ annotate("text", x=1948, y=200000, label="wP")+ annotate("text", x=1998, y=200000, label="aP")
```

Q3) Pertussis has increased in frequency in comparison to before the aP vaccine. This could be explained by aP being less effective than wP.

```{r}
library(jsonlite)
subject <- read_json("https://www.cmi-pb.org/api/subject", simplifyVector = TRUE) 
```

```{r}
head(subject, 3)
```

Q4)96 Q5) 66 Female, 30 Male

```{r}
table(subject$biological_sex)
```

```{r}
table(subject$biological_sex,subject$race)
```
Native,Asian,Black,Multiple,other,not reported,white Female: 0,18,2,8,1,10,27
Male:1,9,0,2,1,4,13

```{r}
library(lubridate)
```
```{r}
subject$age <- today() - ymd(subject$year_of_birth)
library(dplyr)

ap <- subject %>% filter(infancy_vac == "aP")

round( summary( time_length( ap$age, "years" ) ) )
# wP
wp <- subject %>% filter(infancy_vac == "wP")
round( summary( time_length( wp$age, "years" ) ) )
```
Q7) i) 36,ii)25,iii) yes they are significantly different 
Q8) 
```{r}
int <- ymd(subject$date_of_boost) - ymd(subject$year_of_birth)
age_at_boost <- time_length(int, "year")
head(age_at_boost)
```

```{r}
ggplot(subject) +
  aes(time_length(age, "year"),
      fill=as.factor(infancy_vac)) +
  geom_histogram(show.legend=FALSE) +
  facet_wrap(vars(infancy_vac), nrow=2) 
```
Q9) yes they are significantly different
```{r}

specimen <- read_json("https://www.cmi-pb.org/api/specimen", simplifyVector = TRUE) 
titer <- read_json("https://www.cmi-pb.org/api/ab_titer", simplifyVector = TRUE) 
```
Q10)
```{r}
meta <- inner_join(specimen, subject)
dim(meta)
head(meta)
```
```{r}
abdata <- inner_join(titer, meta)
dim(abdata)
```
Q11)

 IgE  IgG IgG1 IgG2 IgG3 IgG4 
6698 1413 6141 6141 6141 6141 
```{r}
table(abdata$isotype)
```
```{r}
table(abdata$visit)
```
There is significantly less people

```{r}
ig1 <- abdata %>% filter(isotype == "IgG1", visit!=8)
head(ig1)
```


```{r}
ggplot(ig1) +
  aes(MFI, antigen) +
  geom_boxplot() + 
  facet_wrap(vars(visit), nrow=2)
```
Q14) FIM2/3 Antigen seems to be responding. Maybe this antigen is what corresponds to pertussis

```{r}
ggplot(ig1) +
  aes(MFI, antigen, col=infancy_vac ) +
  geom_boxplot(show.legend = FALSE) + 
  facet_wrap(vars(visit), nrow=2) +
  theme_bw()
```
Q15)

```{r}
filter(ig1, antigen=="Measles") %>%
  ggplot() +
  aes(MFI, col=infancy_vac) +
  geom_boxplot(show.legend = FALSE) +
  facet_wrap(vars(visit)) +
  theme_bw()
```

```{r}
filter(ig1, antigen=="FIM2/3") %>%
  ggplot() +
  aes(MFI, col=infancy_vac) +
  geom_boxplot(show.legend = FALSE) +
  facet_wrap(vars(visit)) +
  theme_bw()
```
Q16) over time the antigen levels of the aP grow up to visit 4 and seem to decrease gradually with sequential visits

Q17) aP has higher antigen level after the peak visit and seems to almost stay at the peak level compared to the wP that dwindles over time

```{r}
url <- "https://www.cmi-pb.org/api/v2/rnaseq?versioned_ensembl_gene_id=eq.ENSG00000211896.7"

rna <- read_json(url, simplifyVector = TRUE) 
ssrna <- inner_join(rna, meta)
```
Q18)
```{r}
ggplot(ssrna) +
  aes(visit, tpm, group=subject_id) +
  geom_point() +
  geom_line(alpha=0.2)
```
Q19)

It is at maximum level at visit 4

Q20)

Yes the ab titer level shows a peak around visit 5 while the rna seq data shows a peak around visit 4. This is very close. The difference may be that the ab titer may have increased as a repsonse to the expression increase and thus there is a slight delay. 

