---
title: "class05"
author: "Daniel Xu"
format: pdf
---

# Our first ggplot

To use the ggplot2 package I first need to have it installed on my computer.

To install my package we use the `install.packages()` command.

Now can I use it?

```{r}
library(ggplot2)
ggplot()
```
```{r}
mpg
```
Our first plot of displ vs hwy
All ggplot() graphs are made in the same way:

-data+aes+geoms
```{r}
ggplot(mpg)+aes(displ,hwy)+geom_point()
```
I can add more layers:

```{r}
ggplot(mpg)+aes(displ,hwy)+geom_point()+geom_smooth(method=lm, se=TRUE)
```
# Plot of gene expression data
```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```

>Q. How many genes are in this dataset?

```{r}
nrow(genes)
```

What are the colnames?
```{r}
colnames(genes)
```

```{r}
ggplot(genes,aes(Condition1,Condition2))+geom_point()
```

Let's add some color. To do this we can add another aes() mapping of color to the ` State` column in our data'

```{r}
ggplot(genes,aes(Condition1,Condition2,color=State))+geom_point()
```
>Q how many genes are upregulated and downregulated

```{r}
head(genes)

```

To get at just the State column

```{r}
table(genes$State)
```
```{r}
ggplot(genes,aes(Condition1,Condition2,color=State))+geom_point()+scale_colour_manual( values=c("blue","gray","red") )
p<-ggplot(genes,aes(Condition1,Condition2,color=State))+geom_point()+scale_colour_manual( values=c("blue","gray","red") )
```

Adding anotations

```{r}
p+labs(title="Gene Expression Changes Upon Drug Treatment",x="Control (no drug)",y="Drug Treatment",caption="Made by me")

```




