---
title: "class07"
author: "Daniel Xu"
format: html
---
#k means clustering
test

```{r}
x<-rnorm(10)
hist(x)
```
center around -3
```{r}
y<-rnorm(30,-3)
z<-rnorm(30,+3)
temp<-c(y,z)
x<-cbind(temp,rev(temp))
plot(x)
```
kmeans test
```{r}
km<-kmeans(x,centers=3,nstart=20)
km
#how many points are in each cluster
km$size
```
Q what components of your result object details
```{r}
cluster<-km$cluster
km$center
plot(x,col=cluster)
points(km$center,col="blue",pch=15,cex=2)
```

heirarchal test
```{r}
hc<-hclust(dist(x))
plot(hc)
cluster<-cutree(hc,h=4)
c2<-cutree(hc,k=2)
plot(x,col=cluster)
plot(x,col=c2)

```
# Principal Componenet Analysis(PCA)

```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url)
dim(x)
```
Q1)
There are 17 rows with 5 columns

```{r}
rownames(x) <- x[,1]
x<-x[,-1]
head(x)
dim(x)
```
```{r}
x <- read.csv(url, row.names=1)
head(x)
```
Q2) The second method is better because there is less code. It's usually better 
to specify which row is the names rather than use exclusion of the names

```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
barplot(as.matrix(x), col=rainbow(nrow(x)))
```

Q3
You can remove beside=T which results in that plot

Q5
```{r}
pairs(x, col=rainbow(10), pch=16)
```
If they lie on the diagonal that means, that the two countries have similar
consumption of that commodity. The pairwise plot puts each countries on one axis
and plots the values for each commodity with those country axes. 

Q6
North Ireland seems to produce very different amounts when compared to other
countries as many of the commodities stray from the diagonal for every single 
other UK country specifically potatoes and soft drinks.

```{r}
pca<-prcomp(t(x))
summary(pca)
```
Q7
```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x))
```
Q8
```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x),col=rainbow(4))
```
```{r}
v <- round( pca$sdev^2/sum(pca$sdev^2) * 100 )
v
barplot(v, xlab="Principal Component", ylab="Percent Variation")
```

```{r}
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```
```{r}
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,2], las=2 )
```
Q9)
The two groups that are prominent are soft drinks and fresh potatoes.
This tells us that fresh_potatoes pushes N. ireland up for fresh potatoes and 
down for soft drinks. Specifically, N.ireland consumes more fresh potatoes and
consumes less soft drinks compared to other UK countries

```{r}
biplot(pca)
```
```{r}
url2 <- "https://tinyurl.com/expression-CSV"
rna.data <- read.csv(url2, row.names=1)
head(rna.data)
dim(rna.data)
```
Q10
There are 10 genes with 100 samples
