---
title: "class10"
author: "Daniel Xu"
format: pdf
---

```{r}
candy_file <- "candy-data.csv"

candy = read.csv(candy_file, row.names=1)
head(candy)
```
```{r}
dim(candy)
```
Q1) 85
```{r}
sum(candy$fruity>0)
```

Q2)38 
```{r}
candy["Kit Kat", ]$winpercent
candy["Tootsie Roll Snack Bars", ]$winpercent
```


Q3) Kitkat, 76.7686
Q4) 76.7686
Q5) 49.6535

```{r}
library("skimr")
skim(candy)
```
Q6) winpercent
Q7) boolean value, true or false
Q8)
```{r}
hist(candy$winpercent)
```
Q9) No, it is slightly skewed

Q10) It is below 50%
```{r}
mean(candy$winpercent[as.logical(candy$chocolate)])
mean(candy$winpercent[as.logical(candy$fruit)])
```
Q11) On average, chocolate is higher
```{r}
 t.test(candy$winpercent[as.logical(candy$chocolate)],candy$winpercent[as.logical(candy$fruit)])
```

Q12) Yes this is stat significant as p value 2.871e-08 is much lower than alpha level of 0.05
```{r}
head(candy[order(candy$winpercent),], n=5)
```

Q13) The 5 least liked are
Nik L Nip, boston Baked Beans,Chiclets, Super Bubble, and Jawbusters
```{r}
tail(candy[order(candy$winpercent),], n=5)
```

Q14 The 5 most liked are snickers, kit kat, twix, reese miniatures, reese peanut butter cup
```{r}
library(ggplot2)
ggplot(candy) + 
  aes(winpercent,rownames(candy)) +
  geom_bar(stat="identity")
```
```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_bar(stat="identity")
```
```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```
Q17) sixlets
Q18) Starburst
```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = Inf)
```

Q19) Chocolate
```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

Q20) Nik L Lip, Ring Pop, Nestle Smarties,HersheyOs Krackel,HersheyOs Milk Chocolate, Least popular is Nik L Lip

```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```
Q22) 
fruity and chocolate
Q23) any category with itself. If not that, it is winpercent and chocolate 

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```
```{r}
plot(pca$x[,1],pca$x[,2])
```
```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```
```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])
```

```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```
```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = Inf)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```
```{r}
library(plotly)
ggplotly(p)
```
```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```
Q24) Fruity, hard, and pluribus. These make sense as popularity of these type of candy is low
